---
title: "Untitled"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE,
  dpi = 300,
  out.width = "100%",
  fig.align = "center",
  fig.width = 8,
  fig.asp = 0.618, # 1 / phi
  fig.show = "hold",
  dev = "png",
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)

library(tidyverse)
library(sf)
library(crsuggest)

library(ggpmthemes)

theme_set(theme_light_modified(base_family = "Montserrat"))

theme_update(
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  strip.background = element_blank(),
  strip.text = element_text(face = "bold", size = 14)
)
```

## Load data

```{r}
url_predators <- pins::pin(
  "https://cn.dataone.org/cn/v2/resolve/urn:uuid:9ffec04c-7e2d-41dd-9e88-b6c2e8c4375e")

url_nests <- pins::pin("https://cn.dataone.org/cn/v2/resolve/urn:uuid:982bd2fc-4edf-4da7-96ef-0d11b853102d")

predators <- read_csv(url_predators) %>% 
  janitor::clean_names()

nests <- read_csv(url_nests) %>% 
  janitor::clean_names()
```

## Let's make a geographical map

### Convert to spatial object

```{r}
nests_sf <- nests %>% 
  drop_na(long_corrected, lat_corrected) %>% 
  st_as_sf(coords = c("long_corrected", "lat_corrected"), crs = 4326)

crsuggest::suggest_crs(nests_sf)
```

### Select the projection and crop the data

```{r}
nests_sf <- nests_sf %>% 
  st_transform(6351)

map <- rnaturalearth::ne_coastline(returnclass = "sf", scale = "large") %>%
  st_transform(6351) %>%
  st_crop(nests_sf %>% st_buffer(1e4))
```

### Create a map

```{r}
nests_sf %>% 
  ggplot() +
  geom_sf(data = map) +
  geom_sf(aes(color = plot), size = 0.1) +
  scale_color_viridis_d(
    guide = guide_legend(
      override.aes = list(size = 2)
    )
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )
```

## Assign common specie names

The given data only contain the species codes. Would be useful to attach the common species names as well. To do so, I will read the [Species Table and Recommended Band Sizes](https://www.pwrc.usgs.gov/bbl/manual/speclist.cfm). It contains species codes and common names as well.

```{r}
species_lut <- rvest::read_html(
  "https://www.pwrc.usgs.gov/bbl/manual/speclist.cfm"
) %>%
  rvest::html_table() %>%
  .[[1]] %>%
  janitor::clean_names() %>%
  select(species = alpha_code, common_name) %>% 
  mutate(across(where(is.character), tolower))
  
species_lut

nests <- nests %>% 
  left_join(species_lut, by = "species") %>% 
  relocate(common_name, .after = species)

nests

predators <- predators %>% 
  left_join(species_lut, by = "species") %>% 
  relocate(common_name, .after = species)
```

Step #4: Calculate total number of predators by year and species, and plot the result.

```{r}
total_predators_per_year <- predators %>% 
  count(year, species, common_name) %>% 
  mutate(n = case_when(
    is.na(common_name) ~ 0L,
    TRUE ~ n
  )) 

total_predators_per_year

total_predators_per_year %>% 
  drop_na(common_name) %>% 
  ggplot(aes(x = year, y = n, color = common_name)) +
  geom_line(size = 1) +
  geom_point(size = 3, show.legend = FALSE) +
  paletteer::scale_color_paletteer_d(
    "ggthemes::stata_s2color",
    guide = guide_legend(
      override.aes = list(size = 2)
    )
  )
```

Same data, but using a barplot.

```{r}
total_predators_per_year %>% 
  drop_na(common_name) %>% 
  ggplot(aes(x = year, y = n, fill = common_name)) +
  geom_col() +
  paletteer::scale_fill_paletteer_d("ggthemes::stata_s2color")
```

Step #5: Calculate total number of eggs predated by year and species.

```{r}
total_nest_per_year <- nests %>% 
  select(year, site, nest_id, species) %>% 
  count(year, site, name = "total_nest")

total_nest_per_year
```

Step #6: Calculate total number of predators by year, join to summarized egg predation table, and plot the result.

```{r}
total_predators_per_year <- predators %>% 
  count(year, site, name = "total_predator") 

total_predators_per_year
```

Merge nests and predators.

```{r}
total_nest_per_year
total_predators_per_year

df <- inner_join(total_nest_per_year, total_predators_per_year)

df
```

Visualize if the total number of nest and predator co-vary in time.

```{r}
df %>%
  pivot_longer(starts_with("total")) %>% 
  ggplot(aes(x = year, y = value, color = name)) +
  geom_line() +
  geom_point() +
  facet_wrap(~name, ncol = 1)
```

```{r}
df %>%
  ggplot(aes(x = total_predator, y = total_nest)) +
  geom_point() +
  geom_smooth(se = FALSE)
```